---
title: "Predicting the Probability that a Horse Wins a Race in Hong Kong"
author: "Matthew Lewis"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

--
## Import Data

```{r}
# races <- read.csv("races.csv")
# runs <- read.csv("runs.csv")
```

--
## Data Cleaning and Exploration

```{r}
# cleaning races
# races$race_no < as.factor(races$race_no)
# races$surface <- as.factor(races$surface)

# cleaning runs
# library(dplyr)
# runs <- runs %>% select(won, everything()) # move response variable to front
# runs$won <- as.factor(runs$won) # this is our response variable

# runs$trainer_id <- as.factor(runs$trainer_id)
# runs$jockey_id <- as.factor(runs$jockey_id)
```

```{r}
# str(races)
# str(runs)
```

```{r}
# summary(races)
# summary(runs)
```

```{r}
# combine data frames
# hk <- merge(runs, races, by.x = "race_id", by.y = "race_id")
```

```{r}
# write.csv(hk,file="/Users/mlewis/Desktop/School/Winter 2020/TO 414/Project 3/hk.csv", row.names = FALSE)
```

```{r}
hk <- read.csv("hk.csv")
```


Deleting columns which proved to be insignificant.
```{r}
hk$won <- as.factor(hk$won) # this is our response variable

hk$position_sec5 <- NULL
hk$position_sec6 <- NULL

hk$behind_sec5 <- NULL
hk$behind_sec6 <- NULL

hk$time5.x <- NULL
hk$time6.x <- NULL

hk$sec_time5 <- NULL
hk$sec_time6 <- NULL
hk$sec_time7 <- NULL

hk$time5.y <- NULL
hk$time6.y <- NULL

hk$time7 <- NULL

hk$place_combination4 <- NULL

hk$place_dividend4 <- NULL

hk$win_combination2 <- NULL
hk$win_dividend2 <- NULL
hk$horse_gear<-NULL
hk$position_sec4<-NULL
hk$behind_sec4<-NULL
hk$time4.x<-NULL
hk$time4.y<-NULL
hk$sec_time4<-NULL
hk$place_dividend3<-NULL
hk$place_combination3<-NULL


```

```{r}
hk$horse_ratings<-NULL #I dont really know what to do with that column. Maybe to calculte an averages, but how?
```


Overall structure. 

```{r}
str(hk)
```


Its useful to see how many values are missing in each column. It will also help us to make a decision which columns are insignificant. 

```{r}
sapply(hk, function(x) sum(is.na(x)))
```

Useful information can give us how many unique value there are in each column. 
From here we can see there are 6348 races overall. 4405 horses and 186 jockeys competed against each other overall. 

```{r}
sapply(hk, function(x) length(unique(x)))
```

```{r}
table(hk$horse_country)
table(hk$horse_type)
table(hk$going)
table(hk$horse_ratings)
table(hk$config)

```


The next step is to turn as many factors as possible to numeric values. The majority of horses are coming from 4 countries: Australia, New Zeland,Great Britain and USA. So we turned them into the dummy variables and combined other countries into one variable. 

Most of the horses are Gelding type so we combined all other horse types into one value. 

There are 2 tracks represented in the dataset and 6 configurations of the tracks. Since each type represented equally we turned them into its own dummy value. Additional info regarding tracks and its configuration can be found at https://racing.hkjc.com/racing/english/racing-info/racing_course.asp

The condition is also a very important factor on whether the horse is going to win or not. 
  

```{r}
hk$horse_country<-ifelse(hk$horse_country=="AUS",0,
                         ifelse(hk$horse_country=="NZ",1,
                                ifelse(hk$horse_country=="IRE",2,
                                       ifelse(hk$horse_country=="GB",3,
                                              ifelse(hk$horse_country=="USA",4,5)))))

hk$horse_type<-ifelse(hk$horse_type=="Gelding",0,1)

hk$venue<-as.numeric(hk$venue)

hk$config<-ifelse(hk$config=="A",0,
                  ifelse(hk$config=="A+3",1,
                         ifelse(hk$config=="B",2,
                                ifelse(hk$config=="B+2",3,
                                       ifelse(hk$config=="C",4,5)))))


hk$going<-ifelse(hk$going=="FAST",0,
                 ifelse(hk$going=="GOOD",1,
                        ifelse(hk$going=="GOOD TO FIRM",2,
                               ifelse(hk$going=="GOOD TO YIELDING",3,
                                      ifelse(hk$going=="SLOW",4,
                                             ifelse(hk$going=="SOFT",5,
                                                    ifelse(hk$going=="WET FAST",6,
                                                           ifelse(hk$going=="WET SLOW",7,
                                                                  ifelse(hk$going=="YIELDING",8,9)))))))))

```

Since the data is a factor in our dataset, we turned it into data type and calculated which day of the week represents each date. 


```{r}
library(lubridate)
hk$date<-ymd(hk$date)
hk$day<-wday(hk$date)
```

Since we have some missing values in prize column it might be appropriate to remove them with average value.
```{r}
length(which(hk$prize==800000))
hk$prize[is.na(hk$prize)]<-round(mean(hk$prize,na.rm=T))
```



```{r}
str(hk)
```


```{r}
table(is.na(hk))
sum(is.na(hk))
```


Average of place_odds but prof told us to keep the NA's in that case.

```{r}
hk$place_odds[is.na(hk$place_odds)]<-round(mean(hk$place_odds,na.rm=T))
```






#2 I also think we should get rid of horse ratings? Its just a range. Take a mid point of a range!

```{r}


```


```{r}
#levels(hk$horse_country)
#levels(hk$horse_type)
```

Can I convert venue to numeric?
```{r}
#hk1=hk %>% sample_frac(.1)

```


```{r}
str(hk)
summary(hk)
```

--
## Data Visualization



--
## Logistic Regression Models

```{r}
logistic0<-glm(won~.,data=hk,family = "binomial")
summary(logistic0)
```


### Logistic Regression. 
```{r}
logistic1 <- glm(won ~ horse_age + horse_country + horse_type + horse_rating + declared_weight + draw + place_odds + trainer_id + jockey_id, data = hk, family = "binomial")
summary(logistic1)
```
## Predictive Models

### Normalize Data

```{r}
# normalizing data for HK

```

### kNN

```{r}
set.seed(17)
r_hk <- hk[sample(nrow(hk)),]
```



```{r}
r_hk_norm <- as.data.frame(model.matrix(~.-1, r_hk))

normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

r_hk_norm <- as.data.frame(lapply(r_hk_norm[1:50], normalize))
```



```{r}
train_x <- r_hk_norm[1:22000, -ncol(r_hk_norm)]
test_x <- r_hk_norm[22001:27501, -ncol(r_hk_norm)]
train_y <- r_hk_norm[1:22000, ncol(r_hk_norm)]
test_y <- r_hk_norm[22001:27501, ncol(r_hk_norm)]

alpha_train <- r_hk_norm[1:22000, ]
alpha_test <- r_hk_norm[22001:27501, ]
```


```{r}
library(class)
r_hk_norm_pred <- knn(train = train_x, test = test_x, cl = train_y, k = 7) # no missing values are allowed
library(gmodels)
CrossTable(x = test_y, y = r_hk_norm_pred, prop.chisq = FALSE)
```






## Conclusion

**Conclusion**

